% Lydia Lee, modified from EE16A-In-A-Day Fall 2018
\qns{Polarity Check}\ \\
Determine if the following systems are in negative feedback.
\begin{enumerate}
\qitem\ \\
	\begin{center}
		\input{../../questionBank/week08/q_polarity_check_figs/inverting_amp_gnd}
	\end{center}
\ans{
	This is in negative feedback, and in fact is a fairly well-known topology called the inverting amplifier.
	\begin{itemize}
		\item For our initial stimulus, we'll kick $v_\text{out}$ up
		\item How does $u_-$ respond? Well, we can't assume $u_+ = u_-$ (because that requires the system to be in negative feedback), so we'll look at the resistors. Note that this is a voltage divider with $u_-$ in the and, $v_\text{out}$ at the top, so when $v_\text{out}$ moves upward, $u_-$ travels upward too
		\item If $u_-$ goes up, that means $v_\text{out}$ goes down because of the op amp---the opposite direction as the iniital stimulus!
	\end{itemize}
}

\qitem\ \\
	\begin{center}
		\input{../../questionBank/week08/q_polarity_check_figs/blocks}
	\end{center}
\ans{
	This is \textit{not} in negative feedback! Going through the process of checking:
	\begin{itemize}
		\item For our initial stimulus, we'll kick $v_\text{out}$ up
		\item With the $-5$, that means $u_-$ goes down
		\item If $u_-$ goes down, that means $v_\text{out}$ goes up---the same direction as the initial stimulus!
	\end{itemize}
}


\qitem\ \\
	\begin{center}
		\input{../../questionBank/week08/q_polarity_check_figs/tricky}
	\end{center}

\ans{
	This is indeed in negative feedback! Let's go through the process of checking:
	\begin{itemize}
		\item For our initial stimulus, we'll kick $v_\text{out}$ up
		\item With the $-5$, that means $u_+$ goes down
		\item If $u_+$ goes down, that means $v_\text{out}$ goes down---the opposite direction of the initial stimulus!
	\end{itemize}
}
\end{enumerate}