% Author: Mudit Gupta
% Email: mudit@berkeley.edu

% Credits: https://www.mathway.com/examples/algebra/linear-transformations/proving-a-transformation-is-linear?id=266

\qns{Are you linear?}

\begin{enumerate}
\qitem{Consider a matrix $\mathbf{S}$ that transforms a vector $\vec{x} = \begin{bmatrix} a \\ b \\ c\end{bmatrix}$ to $\vec{y} = \begin{bmatrix} a - b - c \\ a-b-c \\ a-b+c\end{bmatrix}$. Note that $a, b, c$ can take on any values in $\mathbb{R}$. In other words, $\mathbf{S}\vec{x} = \vec{y}$. Is this transformation linear?}

\sol{Please note that this might be the first time students are thinking of matrices as transformations. Let this settle in. The fact that a matrix is essentially a function that takes one vector and makes it a different vector. This is no different from a real-valued function like $f(x) = x^2$, except the only difference is that $x$ is a vector, and $f$ is a matrix.}

\ans{
	To prove whether a transformation is linear, we must check whether it preserves scalar multiplication, addition and the zero vector. \\

	\textbf{Scalar multiplication} \\
	Let $\alpha \in \mathbb{R}$. Is $\mathbf{S}(\alpha\vec{x}) = \alpha\vec{y}?$ \\
	$\mathbf{S}\begin{bmatrix}\alpha a \\ \alpha b \\ \alpha c \end{bmatrix} = \alpha \begin{bmatrix} a - b - c \\ a-b-c \\ a-b+c\end{bmatrix}$. Try it! \\ \\

	\textbf{Addition} \\
	Is $\mathbf{S}(\vec{x}_1 + \vec{x}_1) = \mathbf{S}\vec{x}_1 + \mathbf{S}\vec{x}_2?$ \\
	Let $\vec{x}_1 = \begin{bmatrix} a_1 \\ b_1 \\ c_1 \end{bmatrix}$ and $\vec{x}_2 = \begin{bmatrix} a_2 \\ b_2 \\ c_2 \end{bmatrix}$. Then $\mathbf{S}(\vec{x}_1 + \vec{x}_2) = \mathbf{S}\vec{x}_1 + \mathbf{S}\vec{x}_2.$ Try it out!

	\textbf{Zero vector} \\
	Is $\mathbf{S}\vec{0} = \vec{0}$? Yes. \\

	This proves that $\mathbf{S}$ is indeed a linear transformation.
}

\sol{At the end of this, get the students to ask you "well... but... since matrix-vector multiplication is linear, of course every matrix is a linear operator!!" This should be the next question they ask.}

\qitem{
	Now let's consider another matrix $\mathbf{Q}$ which takes a vector $\vec{x} = \begin{bmatrix} a \\ b \\ c \end{bmatrix}$ to $\begin{bmatrix} a + 5 \\ b + c \end{bmatrix}$. Is this matrix a linear operator? 
}

\sol{
	There might be a couple of students who immediately see the answer that the zero vector won't be preserved, but try to make sure they don't just blurt out the answer. Let everyone realize this by themselves.
}

\ans{
	Let's try the preservation of the zero vector first. Is $\mathbf{Q}\vec{0} = \vec{0}?$ Nope, it is $\begin{bmatrix} 5 \\ 0 \\ 0 \end{bmatrix}$. This matrix is not a linear operator! Notice that even though matrix-vector multiplication is generally linear 
}

\qitem{
	Let's dive deeper. Write out the matrix $\mathbf{S}$ and $\mathbf{Q}$. Are they invertible? 
}

\ans{
	$\mathbf{S} = \begin{bmatrix} 1 & -1 & -1 \\ 1 & -1 & -1 \\ 1 & -1 & 1 \end{bmatrix}$. This matrix is not invertible, but it was still linear! \\ \\

	Writing out the matrix for $\mathbf{Q}$ is actually a trick question. There is no easy way to do this. In fact, you cannot simply write it without considering the context in which it is being used. Say that the context is multiplication with a vector. \\
	$$\mathbf{Q}\begin{bmatrix} a \\ b \\ c\end{bmatrix} = \begin{bmatrix} a + 5 \\ b \\ c \end{bmatrix}$$ In this case, $$\mathbf{Q} = \begin{bmatrix} 1 & \dfrac{5}{b} & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{bmatrix}$$ 
	Notice how it is impossible to write the matrix out entirely using just numbers and that we need to use either $a$, $b$ or $c$ inside the matrix itself. 
	Finally, note that the first row of this matrix could be written in other ways too. It could be $\begin{bmatrix} 1 + \dfrac{5}{a} & 0 & 0\end{bmatrix}$, or $\begin{bmatrix} 1 & \dfrac{3}{b} & \dfrac{2}{c}\end{bmatrix}$ too. But the essential idea is that a \textbf{non-linear transformation matrix cannot be expressed using just scalars}. \\
	So it is invertible? We can't say. 
}



\end{enumerate}